---
title: "Integrated Step Selection Functions"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

In this document, we will use integrated step selection analysis (iSSA) to model movement decisions of the Sahtu muskoxen. Models will be fit using the **amt** package. We'll start by loading the data and required packages:

```{r}
#| label: load-packages
#| include: false

library(amt)
library(TwoStepCLogit)
library(glmmTMB)
library(here)
library(sf)
library(terra)
library(tidyverse) 
source(here::here("scripts/functions/hmm_functions.R"))
```

```{r}
#| label: load-data
#| echo: false

musk_collar_filt <- readRDS(here::here("data/processed/musk_collar_filt.rds"))
snowdepth_locs <- readRDS(here::here("data/processed/snowdepth_locs.rds")) %>%
  sf::st_drop_geometry() %>%
  dplyr::select(Id_Number, Latitude, Longitude, datetime, snow_depth)
fire_data_subset <- readRDS(here("data/processed/fire_data_subset.rds"))

```

```{r}
#| label: prep_data
#| dependson: "load-data"

### group data by ID and season
musk_collar_season <- musk_collar_filt %>%
  left_join(snowdepth_locs) %>%
  group_by(year, Id_Number) %>%
  mutate(season = season_delineation(datetime, snow_depth, 0.1),
         year_min = minimum_year_season(year,datetime,season))  %>%
  mutate(id_season = str_c(Id_Number,"_",season)) %>%
  ungroup() 
```

Next let's look at some of the relationships between the spatial data we'll be using in the models. First let's look at historical fire data and how it relates to forest cover data:

```{r}
#| label: fire_hist_data
#| dependson: "prep_data"
#| out-width: 100%
#| layout-ncol: 1

# nbr_proj <- terra::rast(here("data/processed/ntems_fire_nbr_proj.tif"))
# # nbr_cat <- nbr_proj
# # nbr_cat[nbr_proj>66] <- 3
# # nbr_cat[nbr_proj<=66&nbr_proj>27] <- 2
# # nbr_cat[nbr_proj<=27] <- 1
# # plot(nbr_cat)

ggplot() +
  geom_sf(data = musk_collar_season) +
  geom_sf(data = fire_data_subset %>% 
            st_transform(st_crs(musk_collar_season)) %>%
            st_crop(musk_collar_season) %>%
            filter(fireyear<=2012),
          colour = NA,
          aes(fill = fireyear), alpha = 0.8) +
  scale_fill_viridis_c() +
  ggtitle("Fires from 1997 - 2012")

prcC_2020_proj <- terra::rast(here("data/processed/prcC_2020_proj.tif"))
lc_2010_proj <- terra::rast(here("data/processed/lc_2010_proj.tif"))

ggplot() +
  tidyterra::geom_spatraster(data = lc_2010_proj) +
  geom_sf(data = fire_data_subset %>% 
            st_crop(lc_2010_proj) %>%
            filter(fireyear<=2010,
                   fireyear>=1985) %>%
            nngeo::st_remove_holes(),
          aes(colour = fireyear), fill = NA, linewidth = 1) +
  # scale_fill_viridis_c(trans = "sqrt") +
  scale_colour_gradient(high = "darkred", low = "white") +
  labs(fill = "SCANFI canopy\ncover in 2020")
  

```

The first image shows that muskoxen appear to favour areas impacted by recent fires to some degree. The second image shows that canopy cover is also impacted by recent fires, although the relative recency of the fire doesn't seem to have a large effect on differences in canopy cover across areas impacted by fires. In any case, it will be important to account for changes in canopy cover that have occurred since the collars stopped recording as these areas likely had higher canopy cover during the monitoring period.

```{r}
#| label: dem_data
#| dependson: "prep_data"
#| out-width: 100%
#| layout-ncol: 1

mrtri3_proj <- terra::rast(here("data/processed/mrtri3_proj.tif"))

ggplot() +
  tidyterra::geom_spatraster(data = mrtri3_proj) +
  geom_sf(data = musk_collar_season, colour = "red", alpha = 0.1) +
  scale_fill_viridis_c(trans = "sqrt") 

# musk_collar_season %>%
#   leaflet::leaflet() %>%
#   leaflet::addCircles() %>%
#   leaflet::addRasterImage(mrtri3_proj %>%
#                             terra::aggregate(5))
```

There does seem to be avoidance of some areas with high TRI values, however muskox do appear to cross ridge tops to get to potential foraging areas on the other side. Lastly, let's look at the proximity to water and how it relates to muskox movements.

```{r}
#| label: waterprox
#| dependson: "prep_data"
#| out-width: 100%
#| layout-ncol: 1

water_dist <- terra::rast(here("data/processed/water_dist.tif"))

ggplot() +
  tidyterra::geom_spatraster(data = log(water_dist/1000+1)) +
  geom_sf(data = musk_collar_season, colour = "red") 
```

Doesn't seem to be a huge correlation between GPS points and proximity to water, but still worth exploring further in some models.

```{r}
#| label: convert_for_amt
#| dependson: "prep_data"

issa_data <- musk_collar_season %>%
  sf::st_transform(32609) %>%
  mutate(x = sf::st_coordinates(.)[,1],
         y = sf::st_coordinates(.)[,2]) %>%
  sf::st_drop_geometry()


### convert data into tracks for amt package
issa_track <- make_track(issa_data, 
                   x, 
                   y, 
                   datetime, 
                   id_season = id_season,
                   id = Id_Number,
                   season = season,
                   crs = 4326)

### nest tracks by ID season
issa_track_nest <- issa_track %>%
  nest(data = -c(id,id_season,season))

head(issa_track_nest)

```

## ISSA Models

Lets prepare the data for use in ISSA. We'll first resample the data to ensure each step is 8 hours in length. Next we will convert groups of steps into bursts, where a new burst occurs when the time between subsequent steps is greater than 8 hours. Lastly, we will sample random steps for each observed step so that we can compare available and used locations. We use the recommended sample size of 10 for available steps, but will test the robustness of these models by increasing the sample size once we have an appropriate model.

```{r}
#| label: create_bursts
#| dependson: "prep_data"


### create bursts whenever time between steps exceeds 8 hours and
### transform locations into steps
issa_bursts_summer <- issa_track_nest %>%
  filter(season == "summer") %>%
  mutate(
    data_resamp = map(data, function(x)
      x %>% track_resample(rate = hours(8), tolerance = hours(1))
      ),
    steps = map2(data,data_resamp, function(x,y)
      if(length(unique(y$burst_))==1)
        x %>% steps()
      else y %>% steps_by_burst()
      ),
    rand_steps = map(steps, function(x){
      set.seed(5)
      x %>% random_steps(n_control = 10)
      })
  )

issa_bursts_winter <- issa_track_nest %>%
  filter(season == "winter") %>%
  mutate(
    data_resamp = map(data, function(x)
      x %>% track_resample(rate = hours(8), tolerance = hours(1))
      ),
    steps = map2(data,data_resamp, function(x,y)
      if(length(unique(y$burst_))==1)
        x %>% steps()
      else y %>% steps_by_burst()
      ),
    rand_steps = map(steps, function(x){
      set.seed(5)
      x %>% random_steps(n_control = 10)
      })
  )

```

Next we assign covariate data to each step.

```{r}
#| label: assign_covariates
#| dependson: "create_bursts"

prcC_2020_proj <- terra::rast(here("data/processed/prcC_2020_proj.tif"))
mrtri3_proj <- terra::rast(here("data/processed/mrtri3_proj.tif"))
mrdtm_proj <- terra::rast(here("data/processed/mrdtm_proj.tif"))
water_dist <- terra::rast(here("data/processed/water_dist.tif"))
lc_2010_proj <- terra::rast(here("data/processed/lc_2010_proj.tif"))
sm_proj <- terra::rast(here("data/processed/sm_proj.nc"))
nbr_proj <- terra::rast(here("data/processed/ntems_fire_nbr_proj.tif"))
names(nbr_proj) <- "nbr"
rasts <- list.files(here("data/processed/fire_year"), 
                    pattern = "*.tif$", 
                    full.names = TRUE)
fire_year <- rast(rasts)

issa_data_summer <- issa_bursts_summer %>%
  mutate(
    data_covs = map2(rand_steps, id, function(x,y){
      set.seed(5)
      x  %>%
        mutate(cos_ta = cos(ta_),
               log_sl = log(sl_),
               id = y) %>%
        extract_covariates(prcC_2020_proj, where = "both") %>%
        extract_covariates(mrtri3_proj) %>%
        extract_covariates(mrdtm_proj, where = "both") %>%
        extract_covariates(water_dist) %>%
        extract_covariates(lc_2010_proj) %>%
        extract_covariates(nbr_proj) %>%
        extract_covariates_var_time(fire_year, max_time = years(1),
                                    when = "before", name_covar = "fire_year") %>%
        drop_na(ta_)
      })
  )

issa_data_winter <- issa_bursts_winter %>%
  mutate(
    data_covs = map2(rand_steps, id, function(x,y){
      set.seed(5)
      x  %>%
        mutate(cos_ta = cos(ta_),
               log_sl = log(sl_),
               id = y) %>%
        extract_covariates(prcC_2020_proj, where = "both") %>%
        extract_covariates(mrtri3_proj) %>%
        extract_covariates(mrdtm_proj, where = "both") %>%
        extract_covariates(water_dist) %>%
        extract_covariates(lc_2010_proj) %>%
        extract_covariates(nbr_proj) %>%
        extract_covariates_var_time(fire_year, max_time = years(1),
                                    name_covar = "fire_year") %>%
        extract_covariates_var_time(sm_proj, max_time = days(2),
                                    name_covar = "snow_depth",
                                    where = "start") %>%
        drop_na(ta_)
      })
  )
```

Next we will collapse each muskox-specific dataset into a combined dataset for each season. We will assign a new ID to each muskox burst and calculate some new covariates for the models. We will also examine the covariates for any correlation. \

```{r}
#| label: combined_data
#| dependson: "assign_covariates"

issa_data_summer_comb <- list_rbind(issa_data_summer$data_covs) %>%
  group_by(id, step_id_) %>%
  mutate(strat_id = cur_group_id()) %>%
  ungroup() %>%
  mutate(log_water_dist = log(layer/30+1),
         water = ifelse(cover == "Water",1,0),
         open_land = ifelse(cover %in% 
                              c("Temperate or sub-polar shrubland",
                                "Temperate or sub-polar grassland")
                            ,1,0),
         forest = ifelse(cover %in% 
                              c("Temperate or sub-polar needleleaf forest",
                                "Sub-polar taiga needleleaf forest",
                                "Temperate or sub-polar broadleaf deciduous forest",
                                "Mixed forest")
                            ,1,0),
         nbr = replace_na(nbr,0),
         nbr_low = ifelse(is.na(nbr)|nbr<=27,1,0),
         nbr_moderate = ifelse(nbr>27&nbr<=66,1,0),
         nbr_high = ifelse(nbr>66,1,0),
         fire_10 = ifelse(fire_year<=10,1,0),
         fire_20 = ifelse(fire_year>10&fire_year<=20,1,0),
         fire_30 = ifelse(fire_year>20&fire_year<=30,1,0),
         fire_40 = ifelse(fire_year>30&fire_year<=40,1,0),
         fire_10_low = fire_10*nbr_low,
         fire_10_mod = fire_10*nbr_moderate,
         fire_10_high = fire_10*nbr_high,
         fire_20_low = fire_20*nbr_low,
         fire_20_mod = fire_20*nbr_moderate,
         fire_20_high = fire_20*nbr_high)

issa_data_summer_comb %>%
  select(log_sl, cos_ta, scanfi_prcC_end, mrdtm,
         fire_year, open_land, forest) %>%
  cor()

issa_data_winter_comb <- list_rbind(issa_data_winter$data_covs) %>%
  group_by(id, step_id_) %>%
  mutate(strat_id = cur_group_id()) %>%
  ungroup() %>%
  mutate(log_water_dist = log(layer/100+1),
         ### only 3-4 potentital steps in fires aged <10 years
         ### also not enough 1-20 year fires for separate groups based on nbr
         fire_10_20 = ifelse(fire_year<=20,1,0),
         fire_30 = ifelse(fire_year>20&fire_year<=30,1,0),
         fire_40 = ifelse(fire_year>30&fire_year<=40,1,0),
         dtm_diff = mrdtm_end - mrdtm_start,
         open_land = ifelse(cover %in% 
                              c("Temperate or sub-polar shrubland",
                                "Temperate or sub-polar grassland")
                            ,1,0),
         forest = ifelse(cover %in% 
                              c("Temperate or sub-polar needleleaf forest",
                                "Sub-polar taiga needleleaf forest",
                                "Temperate or sub-polar broadleaf deciduous forest",
                                "Mixed forest")
                            ,1,0))

issa_data_winter_comb %>%
  select(log_sl, cos_ta, scanfi_prcC_end, mrdtm, 
         fire_year, open_land, 
         forest, snow_depth) %>%
  cor()


```

Lastly, we will fit models for the summer and winter datasets separately. Models are fit using the ISSA method that allows for random effects. This way we can automatically pool estimates across individuals to come up with a population-level estimate for each covariate.

```{r}
#| label: summer_models
#| dependson: "combined_data"

# m0_summer_2step <- Ts.estim(case_ ~ scale(log_sl)*scale(cos_ta) + 
#                               scale(scanfi_prcC_end) +
#                               scale(mrdtm) +
#                               # scale(dtm_diff) +
#                               scale(log_water_dist) +
#                               open_land + 
#                               scale(log_fire_year) +
#                               strata(strat_id) +
#                               cluster(id),
#                             data = issa_data_summer_comb)
# m0_summer_2step

m0_summer_glmmfit <- glmmTMB(case_ ~ -1 + scale(log_sl)*scale(cos_ta) + 
                            scale(scanfi_prcC_end) +
                            scale(mrdtm) +
                            # scale(dtm_diff) +
                            scale(log_water_dist) +
                            water +
                            fire_10_low +
                            fire_10_mod +
                            fire_10_high +
                            fire_20_low +
                            fire_20_mod +
                            fire_20_high +
                            fire_30 +
                            fire_40 +
                            (1|strat_id) +
                            (0 + scale(log_sl)|id) +
                            (0 + scale(cos_ta)|id) +
                            (0 + scale(log_sl):scale(cos_ta)|id) +
                            (0 + scale(scanfi_prcC_end)|id) +
                            (0 + scale(mrdtm)|id) +
                            # (0 + scale(dtm_diff)|id) +
                            (0 + scale(log_water_dist)|id) +
                            (0 + water|id) +
                            (0 + fire_10_low|id) +
                            (0 + fire_10_mod|id) +
                            (0 + fire_10_high|id) +
                            (0 + fire_20_low|id) +
                            (0 + fire_20_mod|id) +
                            (0 + fire_20_high|id) +
                            (0 + fire_30|id) +
                            (0 + fire_40|id), 
                          family = poisson,
                          data = issa_data_summer_comb,
                          map=list(theta=factor(c(NA,1:15))),
                          start=list(theta=c(log(1e3),rep(0,15))))
summary(m0_summer_glmmfit)

saveRDS(issa_data_summer_comb, here("data/processed/issa_data_summer_comb.rds"))
saveRDS(m0_summer_glmmfit, here("output/m0_summer_glmmfit.rds"))


```

```{r}
#| label: winter_models
#| dependson: "combined_data"

# m0_winter_2step <- Ts.estim(case_ ~ scale(log_sl)*scale(cos_ta) + 
#                               scale(scanfi_prcC_end) +
#                               scale(mrdtm) +
#                               scale(dtm_diff) +
#                               scale(log_water_dist) +
#                               open_land + 
#                               scale(snow_depth):scale(log_sl) +
#                               scale(snow_depth):scale(dtm_diff) +
#                               # scale(log_fire_year) +
#                               strata(strat_id) +
#                               cluster(id),
#                             data = issa_data_winter_comb)
# m0_winter_2step

m0_winter_glmmfit <- glmmTMB(case_ ~ -1 + scale(log_sl)*scale(cos_ta) + 
                            scale(scanfi_prcC_end) +
                            scale(mrdtm) +
                            # scale(dtm_diff) +
                            scale(log_water_dist) +
                            # open_land +
                            fire_10_20 +
                            fire_30 +
                            fire_40 +
                            scale(snow_depth):scale(log_sl) +
                            (1|strat_id) +
                            (0 + scale(log_sl)|id) +
                            (0 + scale(cos_ta)|id) +
                            (0 + scale(log_sl):scale(cos_ta)|id) +
                            (0 + scale(scanfi_prcC_end)|id) +
                            (0 + scale(mrdtm)|id) +
                            # (0 + scale(dtm_diff)|id) +
                            (0 + scale(log_water_dist)|id) +
                            # (0 + open_land|id) +
                            (0 + scale(snow_depth):scale(log_sl)|id) +
                            (0 + fire_10_20|id) +
                            (0 + fire_30|id) +
                            (0 + fire_40|id), 
                          family = poisson,
                          data = issa_data_winter_comb,
                          map=list(theta=factor(c(NA,1:10))),
                          start=list(theta=c(log(1e3),rep(0,10))))

summary(m0_winter_glmmfit)

saveRDS(issa_data_winter_comb, here("data/processed/issa_data_winter_comb.rds"))
saveRDS(m0_winter_glmmfit, here("output/m0_winter_glmmfit.rds"))

```
